\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual Multi-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
some edits on loops

The color factor
$C_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $N_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

In this form the color factors $C_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
C_i+C_j+C_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}. The duality between color and kinematics is the statement that that there exists a way to rearrange factors between
the various kinematic numerators $N_\Gamma$ such that the numerators
obey the same Jacobi identities as the color factors
\begin{equation}
N_i +N_j +N_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.
When the kinematic algebra is not known explicitly, the kinematic numerators are typically constructed from an ansatz.
As an example, the result of such a calculation for four-point NLSM pion scattering at tree level produces the numerators
\begin{align}
&N_s = s(t-u)\\
&N_t = t(u-s)\\
&N_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels so that the net result
is a local function, just as one would expect for the pion tree
amplitude.

@@

In fact, we can make the failure extremely precise.  Start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}:
the double-box, crossed-box, and penta-triangle.  For each of the
three diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop).  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2761 free parameters on the double-box, 2576 free parameters on the
crossed-box, and 5040 free paramters on the penta-triangle.  Then
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  N \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  N\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{eBox}
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
\end{eBox}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut.

\section{Conclusions and Outlook}\label{conclusions}
There is overwhelming evidence that YM and NLSM obey color-kinematics
duality at loop level, even though naive Lagrangians for these theories
do not manifest the duality.  At tree level, the duality can be
manifested by breaking manifest Bose symmetry at the Lagrangian level,
either by explicitly picking different gluon states (as in SDYM) or by
picking states that only restore Bose symmetry in the final amplitude
(as in NLSM).  All of these theories can be encapsulated in the
semi-abelian YM theory presented in \cref{eq:semiYM} which obscures
the Bose symmetry of the external states due to the presence of the
two gauge fields $A_\mu$ and $\bar{A}_\mu$.  The kinetic mixing
between the two gauge fields permits globally color-dual 1-loop
integrands, but special care must be taken so that the integrands
produced from semi-abelian YM correspond to SDYM or NLSM.  Due to the
kinetic mixing that stems from the broken Bose symmetry, it is
impossible to construct a two-loop integrand for semi-abelian YM theory,
that is, the theory is one-loop exact.  Obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry and in this paper it is argued that off-shell locality must
be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands.}

@@

\subsection{Non-local construction of scattering amplitudes}
\label{nonLocalScattering}
For example, consider the simple four-point example of a color-dual
representation of Yang-Mills theory.
\begin{equation}
N_{(12|34)}^{\text{YM}} = \frac{t_8F^4}{3} \frac{s_{13}-s_{12}}{s_{23}{s_{13}}}
\end{equation}
By construction, this functionally symmetric numerator is
anti-symmetric and obeys the Jacobi identity:
\begin{equation}
N_{(12|34)}^{\text{YM}} + N_{(13|42)}^{\text{YM}} +N_{(14|23)}^{\text{YM}} = 0
\end{equation}
However, it must also factorize to kinematics that are consistent with
the local Feynman rules of \cref{eq:ym-lag}. Evaluating the residue on
the $s=s_{12} \to 0$ pole we find:
\begin{align}
\text{Res}\left[\frac{ N_{s}^{\text{YM}} }{s}\right]^{s=0}&= \frac{2}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\\
\text{Res}\left[\frac{N_{t}^{\text{YM}} }{t}\right]^{s=0}&= -\frac{1}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\\
\text{Res}\left[\frac{N_{u}^{\text{YM}} }{u}\right]^{s=0}&= -\frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\end{align}
with $t=s_{23}$ and $u=s_{13}$.
Plugging these into a cubic graph representation of Yang-Mills, the
tree level version of \cref{eq:gen-amp}, we find:
\begin{equation}
  \text{Res}\left[\mathcal{A}_4^{\text{YM}}\right]^{s=0}
  = \frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} (2c_s-c_t-c_u)
  =  \mathcal{A}_{(12|l)}^{\text{YM}}\mathcal{A}_{(-l|34)}^{\text{YM}} 
\end{equation}
where we have applied the color structure Jacobi identity,
$c_s+c_t+c_u=0$. While we can absolutely apply a generalized gauge
trasformation \cite{BCJ} to restore field theoretic locality to the
cubic numerators, we can see clearly that this is an aesthetic
choice. Moreover, by abandoning traditional notions of off-shell
locality, we incidentally gained enough freedom to massage the
color-dual numerators into a form that is manifestly gauge invariant
for all particles. As an organizational principle, pulling out overall
gauge-invariant factors comes with the added advantage of simplifying
a loop-level construction of color-dual numerators for Yang-Mills.

\subsection{Future Directions}\label{sec:Discussion}

Much of the complexity for $D$-dimensional vector theories at
loop-level arises due to the mixing between external polarization and
internal loop momenta. As noted, one advantage of expanding ansatz
construction towards functions of rational kinematics is the prospect
of stripping out overall factors of gauge invariant tensors. Rather
than building an ansatz from the irreducible scalar products of
\cref{eq:ym-basis},
\begin{equation}
N^{\text{YM}}_{m,L} = \sum_i \mathcal{O}_i \mathcal{R}^{(i)}_{m,L}
\end{equation}
where $\mathcal{O}_i$ are the basis tensors of
\cite{Bern:2017tuc,Carrasco:2019yyn}, and $ \mathcal{R}^{(i)}_{m,L}$ is a
rational function of scalar products.

@@

Before proceeding, we note a strange property of the $n$-gon numerator for the pions. In this form, Jacobi relations produce \textit{non-vanishing} values for bubbles on external legs (BELs). However, it is easy to see that the contribution integrates to zero after applying IBP relations/tensor reduction on the bubble. The BEL diagram can be reconstructed from the $n$-gon as follows:
\begin{equation}
N^{\text{BEL}}_{1|2,34} = N^{\text{NLSM}}_{(1234,\ell)}-N^{\text{NLSM}}_{(1243,\ell)}-N^{\text{NLSM}}_{(1342,\ell)}+N^{\text{NLSM}}_{(1432,\ell)}
\end{equation}
where we define the loop momentum to be in between the left most and right most leg on the box. Plugging in particular values for $\ell_i$, we obtain the following expression for the BEL
\begin{equation}
N^{\text{BEL}}_{1|2,34} = s_{12} (\ell+k_1)^2 \ell^{\mu} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} 
\end{equation}
Notice there is an overall factor that cancels one of the propagators. Plugging this in, produces an integral of the following form
\begin{equation}
\mathcal{I}^{\text{BEL}}_{1|2,34} = s_{12} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \int \frac{d^D \ell}{i\pi^{D/2}} \frac{\ell^\mu }{\ell^2-\mu^2} \sim   s_{12}(s_{13}-s_{14}) (\mu^2)^{D/2}
\end{equation}
where we have introduced a mass regulator that will be proportional the the on-shell momentum inside the BEL, $\mu^2 \equiv k_1^2$. Thus, in large enough dimension, this integral suppresses the $\mu^{-2}$ divergence appearing in the denominator of the BEL diagram.

>>>>

The color factor
$C_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,Carrasco:2022jxn,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $N_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors. In this 
form the color factors $C_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
C_i+C_j+C_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}. The duality between color and kinematics is the statement that that there exists a way to rearrange factors between
the various kinematic numerators $N_\Gamma$ such that the numerators
obey the same Jacobi identities as the color factors
\begin{equation}
N_i +N_j +N_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.
When the kinematic algebra is not known explicitly, the kinematic numerators are typically constructed from an ansatz.
As an example, the result of such a calculation for four-point NLSM pion scattering at tree level produces the numerators
\begin{align}
&N_s = s(t-u)\\
&N_t = t(u-s)\\
&N_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels so that the net result
is a local function, just as one would expect for the pion tree
amplitude.

@@

In fact, we can make the failure extremely precise.  Start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}:
the double-box, crossed-box, and penta-triangle.  For each of the
three diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop).  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2761 free parameters on the double-box, 2576 free parameters on the
crossed-box, and 5040 free paramters on the penta-triangle.  Then
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  N \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  N\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{eBox}
\begin{equation}\label{chefsKiss}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
\end{eBox}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut.

\section{Conclusions and Outlook}\label{conclusions}

There is overwhelming evidence that YM and NLSM obey color-kinematics
duality at loop level, even though naive Lagrangians for these theories
do not manifest the duality.  At tree level, the duality can be
manifested by breaking manifest Bose symmetry at the Lagrangian level,
either by explicitly picking different gluon states (as in SDYM) or by
picking states that only restore Bose symmetry in the final amplitude
(as in NLSM).  All of these theories can be encapsulated in the
semi-abelian YM theory presented in \cref{eq:semiYM} which obscures
the Bose symmetry of the external states due to the presence of the
two gauge fields $A_\mu$ and $\bar{A}_\mu$.  The kinetic mixing
between the two gauge fields permits globally color-dual 1-loop
integrands, but special care must be taken so that the integrands
produced from semi-abelian YM correspond to SDYM or NLSM.  Due to the
kinetic mixing that stems from the broken Bose symmetry, it is
impossible to construct a two-loop integrand for semi-abelian YM theory,
that is, the theory is one-loop exact.  Obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry and in this paper it is argued that off-shell locality must
be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands.}

@@

\subsection{Non-local construction of scattering amplitudes}
\label{nonLocalScattering}
As an exemplar of an off-shell non-local construction, consider the simple four-point example of a color-dual
representation of Yang-Mills theory. We can define a functional numerator, $N_{(12|34)}^{\text{YM}}$, as follows, 
\begin{equation}
N_{(12|34)}^{\text{YM}} = \frac{t_8F^4}{3} \frac{s_{13}-s_{12}}{s_{23}{s_{13}}}
\end{equation}
where $N_s = N_{(12|34)}$, $N_t = N_{(14|23)}$ and $N_u = N_{(13|42)}$. By construction, this functionally symmetric numerator is
anti-symmetric and obeys the Jacobi identity:
\begin{equation}
N_{(12|34)}^{\text{YM}} + N_{(13|42)}^{\text{YM}} +N_{(14|23)}^{\text{YM}} = 0
\end{equation}
However, it must also factorize to kinematics that are consistent with
the local Feynman rules of \cref{eq:ym-lag}. Evaluating the residue on
the $s=s_{12} \to 0$ pole we find:
\begin{align}
\text{Res}\left[\frac{ N_{s}^{\text{YM}} }{s}\right]^{s=0}&= \frac{2}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\\
\text{Res}\left[\frac{N_{t}^{\text{YM}} }{t}\right]^{s=0}&= -\frac{1}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\\
\text{Res}\left[\frac{N_{u}^{\text{YM}} }{u}\right]^{s=0}&= -\frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\end{align}
with $t=s_{23}$ and $u=s_{13}$.
Plugging these into a cubic graph representation of Yang-Mills, the
tree level version of \cref{eq:gen-amp}, we find:
\begin{equation}
  \text{Res}\left[\mathcal{A}_4^{\text{YM}}\right]^{s=0}
  = \frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} (2c_s-c_t-c_u)
  =  \mathcal{A}_{(12|l)}^{\text{YM}}\mathcal{A}_{(-l|34)}^{\text{YM}} 
\end{equation}
where we have applied the color structure Jacobi identity,
$c_s+c_t+c_u=0$. In light of our findings in \cref{chefsKiss},  While we can absolutely apply a generalized gauge
transformation \cite{BCJ} to restore field theoretic locality to the
cubic numerators, it is maybe aesthetic
choice.

Moreover, by abandoning traditional notions of off-shell
locality, we incidentally gained enough freedom to massage the
color-dual numerators into a form that is manifestly gauge invariant
for all particles. As an organizational principle, pulling out overall
gauge-invariant factors comes with the added advantage of simplifying
a loop-level construction of color-dual numerators for Yang-Mills.

Much of the complexity for $D$-dimensional vector theories at
loop-level arises due to the mixing between external polarization and
internal loop momenta. As noted, one advantage of expanding ansatz
construction towards functions of rational kinematics is the prospect
of stripping out overall factors of gauge invariant tensors. Rather
than building an ansatz from the irreducible scalar products of
\cref{eq:ym-basis},
\begin{equation}
N^{\text{YM}}_{m,L} = \sum_i \mathcal{O}_i \mathcal{R}^{(i)}_{m,L}
\end{equation}
where $\mathcal{O}_i$ are the basis tensors of
\cite{Bern:2017tuc,Carrasco:2019yyn}, and $ \mathcal{R}^{(i)}_{m,L}$ is a
rational function of scalar products.
\subsection{Future Directions}\label{sec:Discussion}

@@

Before proceeding, we note a strange property of the $n$-gon numerator for the pions. In this form, Jacobi relations produce \textit{non-vanishing} values for bubbles on external legs (BELs). However, it is easy to see that the contribution integrates to zero after applying IBP relations/tensor reduction on the bubble. The BEL diagram can be reconstructed from the $n$-gon as follows:
\begin{equation}
N^{\text{BEL}}_{1|2,34} = N^{\text{NLSM}}_{(1234,\ell)}-N^{\text{NLSM}}_{(1243,\ell)}-N^{\text{NLSM}}_{(1342,\ell)}+N^{\text{NLSM}}_{(1432,\ell)}
\end{equation}
where we define the loop momentum to be in between the left most and right most leg on the box. Plugging in particular values for $\ell_i$, we obtain the following expression for the BEL
\begin{equation}
N^{\text{BEL}}_{1|2,34} = s_{12} (\ell+k_1)^2 \ell^{\mu} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} 
\end{equation}
Notice there is an overall factor that cancels one of the propagators. Plugging this in, produces an integral of the following form
\begin{equation}
\mathcal{I}^{\text{BEL}}_{1|2,34} = s_{12} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \int \frac{d^D \ell}{i\pi^{D/2}} \frac{\ell^\mu }{\ell^2-\mu^2} \sim   s_{12}(s_{13}-s_{14}) (\mu^2)^{D/2}
\end{equation}
where we have introduced a mass regulator that will be proportional the the on-shell momentum inside the BEL, $\mu^2 \equiv k_1^2$. Thus, in sufficiently large dimension, $D>2$, this integral suppresses the $\mu^{-2}$ divergence appearing in the denominator of the BEL diagram.
